---
title: "Determinants Influencing the Sex Ratio of Turtle Offspring"
author: "Yuechen Huang; Shiqi Zheng"
date: "Fall 2023"
output:
  html_document:
    df_print: paged
---

## Set up and Data Wrangling
```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(rotl)

Cleaned_Data <- read.csv(here("Processed/Cleaned_Data.csv"), 
                         stringsAsFactors = TRUE)
theme_yh <- theme_classic(base_size = 14) + 
  theme(axis.text = element_text(color = "black"),
  legend.position = "right",
  legend.text = element_text(size = 10),
  legend.title = element_text(size = 11),
  plot.title = element_text(size = rel(1.5)),
  axis.text.x = element_text(angle = 75, vjust = 1, hjust=1))
```

## Explore the species 
### Phylogenetic tree
We used a "rotl" package to interact with the Open Tree of Life data (https://doi.org/10.1111/2041-210X.12593). First, we need to change the name of the species into something compatible with the package.

```{r generate phylogenetic tree, fig.height=10, fig.width=6, warning=FALSE}
Cleaned_Data$Species_names <- as.character(Cleaned_Data$Species)  # Convert factors to characters

# Replace underscores with spaces
Cleaned_Data$Species_names <- gsub("_", " ", Cleaned_Data$Species_names)

# Create a data frame that calculate the mean sex ratio of each species
Data_species <- Cleaned_Data %>% group_by(Species_names) %>%
  summarise(mean_sex_ratio = mean(Sex_ratio, na.rm = TRUE),
            SD_sex_ratio = sd(Sex_ratio, na.rm = TRUE),
            mean_humidity_sp = mean(Humidity, na.rm = TRUE),
            SD_humidity_sp = sd(Humidity, na.rm = TRUE),
            mean_temp_sp = mean(Mean_Temp, na.rm = TRUE),
            SD_temp_sp = sd(Mean_Temp, na.rm = TRUE)) 
write.csv(Data_species, file = here("Processed/Data_Species.csv"), row.names = FALSE) # Data regroup by species

Data_species$taxa <- rotl::tnrs_match_names(Data_species$Species_names)
tree <- tol_induced_subtree(ott_ids = ott_id(Data_species$taxa))
collaped_tree <- ape::ladderize(ape::collapse.singles(tree, root.edge = FALSE)) ## tried to make this less messy but doesn't work
plot(collaped_tree, cex = .5, label.offset = .1, no.margin = TRUE)
```

### Tree map
Because there are too many species, the plot looks very messy...Then we can try to make a treemap instead (https://r-graph-gallery.com/236-custom-your-treemap.html).

```{r treemap}
library(ggplot2)
## install.packages("treemap")
# install.packages("remotes")
# remotes::install_github("d3treeR/d3treeR")
library(treemap)
library(d3treeR)

species_count <- count(Cleaned_Data, Order, Family, Genus, Species_names)
write.csv(species_count, file = here("Processed/Species_Count.csv"), row.names = FALSE) 

phytree <- treemap(species_count, 
        index = c("Order", "Family", "Genus", "Species_names"),
        vSize = "n",
        type = "index",
        vColor = "index",
        title="Species Occurance in the database",
        palette="Dark2",
        fontsize.labels=c(15,12,11,10),
        fontcolor.labels = c("White","Black","Grey","#ffce86"),
        fontface.labels=c(2,1,1,3),  
        bg.labels=c("transparent"), 
        overlap.labels= 0.15,
        inflate.labels= T,
        )

interactive_tree <- d3tree2(phytree, 
                            rootname = "Species Occurance in the database")
interactive_tree

# save the widget
library(htmlwidgets)
saveWidget(interactive_tree, file=paste0(here(), "/Processed/interactiveTreemap.html"))
```

From the tree maps (ggplot map and interactive tree map) we can see that all of the species are within Testudines Order. The most abundant family is Cheloniidae, Emydidae and Chelydridae. The species we have most in our database is Common snapping turtle (Chelydra serpentina) and Green sea turtle (Chelonia mydas) with a occurrence of 1123 and 881 records.

### Basic statistics about species
We will explore sex ratio and temperature data for different species respectively.

#### Sex ratio among species
First, we need to filter the data because we have too many species. Let's only focus on species that have more than 100 records in our data base
```{r stats}

filtered_sp_df <- filter(species_count, n >= 100)
Sp_100 <- filtered_sp_df$Species_names
Sp_100
sp_100_sp <- filter(Data_species, Species_names
  %in% Sp_100) # Data regroup by species and only species that have 100 records are included
sp_100_data <- filter(Cleaned_Data, Species_names
  %in% Sp_100) # Whole cleaned data and only species that have more than 100 records are included
  
#ggplot(Data_species, aes(x = ))
```

Even without graph, we can see that all of the species have a very high sex ratio (infinite number of sex ratio). This may due to the fact that if we have more records of a specific species, the chance to have extreme high (even infinite) sex ratio records can get higher. Perhaps we need to omit infinite data? 

#### Temperature among families

```{r temp}
ggplot(sp_100_data, aes(x = Species_names, y = Mean_Temp, 
                      fill = Species_names, 
                      label = Mean_Temp)) + 
  geom_boxplot(na.rm = TRUE) + theme_yh +
  scale_fill_brewer(palette="Set3") + 
  labs(x = "Species Names", y = expression("Hatching Temperature ("*~degree*C*")")) 
  # + geom_text(aes(label = round(mean_temp_sp, 2)), vjust = -0.5)  
```
Hatching temperature among all the species are around 25 - 35 degree Celsius. Emys orbicularis have a high variance of hatching temperature for different individuals or communities.

\newpage
## Captive vs Wild turtles
We want to explore the sex ratio of captive and wild turtles among all different species. (Leave it there because of sex ratio problem)
```{r captivity}

```

